MG7: Configurable and scalable 16S data analysis

### 2016-07-26 docs/mg7-paper/out/paper.tex
remove older LCA text, explain why LCA a bit

For each read, first, we select the BEST BLAST HITs (by default 10 Hits)
over a threshold of similarity. To evaluate similarity for this first
filtering of hits we use the Expect value (by default
\(evalue \leq e^{-15}\)) that describes the number of hits one can
``expect'' to see by chance when searching a database of a particular
size. In a second filtering step we filtering those hits that are not
sufficiently good comparing them with the best one. We select the best
HSP (High Similarity Pair) per reference sequence and then choose the
best HSP (that with lowest E-value) between all the selected ones. The
bitscore of this best HSP (called S) is used as reference to filter the
rest of HSPs. All the HSPs with bitscore below the product \(p S\) are
filtered. p is a coefficient fixed by the user to define the bitscore
required, e.g.~if \(p=0.9\) and \(S=700\) the required bitscore
threshold would be \(630\). Once we have the definitive HSPs selected,
we obtain their corresponding taxonomic nodes using the taxonomic
assignments that NCBI provides for all the nt database sequences. Now we
have to analyze the topological distribution of these nodes in the
taxonomy tree: i. If all the nodes forms a line in the taxonomy tree
(are located in a not branched lineage to the tree root) we should
choose the most specific taxID as the final assignment for that read. We
call to this kind of assignment the `in line' exception (see Figure 2
right panel). ii. If not, we should search for the \emph{sensu stricto}
Lowest Common Ancestor (LCA) of all the selected taxonomic nodes (See
Figure 2 left panel). In this approach we decided to use the bitscore
for evaluating the similarity because it is a value that increases when
similarity is higher and depends a lot on the length of the HSP. Some
reads could not find sequences with enough similarity in the database
and then they would be classified as reads with no hits. Advanced
metagenomics analysis approaches \citep{huson2012microbial} have adopted
LCA-based assignment algorithms because it provides fine and trusted
taxonomical assignment.

>>>>

The LCA assignment simply computes the lowest common ancestor of the set
of best hits. Other metagenomics analysis approaches
\citep{huson2012microbial} have adopted LCA-based assignment algorithms,
as it provides the most precise assignment for a set of taxa, with no
further information available\footnote{the lowest common ancestor of a
  set of nodes in a tree is just their product in the tree viewed as a
  poset, with the ``parent of'' partial order}; as our reference
database is already heavily preprocessed and checked for internal
consistency, there is no need for altering the lowest common ancestor in
any way, as is customarily done in other cases.
